### 2024-04-08 README.Rmd
1. Added functions to start simple simulations more easily 2. Added vignette 3. Sara is now package maintainer

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# MethEvolSIM

<!-- badges: start -->
<!-- badges: end -->

Software for simulation of DNA methylation dynamics on different genomic structures under the basic model for the evolution of DNA methylation along genealogies. The methylation state of each CpG dinucleotide is represented as 1 for unmethylated, 2 for partially methylated and 3 for methylated.

# Installation

You can use the development version of MethEvolSIM from git. The development is being implemented with the `devtools` package. It can be loaded and checked using the following:

```{r}
library(devtools)
load_all()
```

`check()` runs the unit tests. The unit tests that test the functions used for writing output files:
- *validationResults.txt*: Is generated in case an error relative to the characteristics of the objects happens and contains both the error and the object.

# Manual of Use

## Structural distribution of CpG sites

The initial spatial distribution of CpG sites is given by the user. A data frame with 3 columns named "start", "end" and "globalState" encodes the structural information. Each row provides the information of a different structure. The number of CpG positions is given by $end - start +1$, and each row start value needs to be as the previous row end value + 1. The globalState encodes island and non-island structures with the character value "U" for islands and "M" for non-islands.

```{r}
# Example 3 structures of length 100: non-island, island, non-island
infoStr <- data.frame(start = c(1, 101, 201),
                        end = c(100, 200, 300),
                        globalState = c("M", "U", "M"))
infoStr
```

```{r}
# Example 2 islands: first of length 1, second of lenght 15
infoStr <- data.frame(start = c(1, 2),
                        end = c(1, 16),
                        globalState = c("U", "U"))
infoStr
```

## Initial methylation state distribution

Can be sampled as specified in the section "Model parameters" when infoStr has only start, end and globalState columns,
or given by the user as additional columns of the data frame. Note that when given, each of the structures frequencies needs to sum up to 1.

```{r}
# Example 3 structures of length 100 with customized initial methylation state equilibrium frequencies
infoStr <- data.frame(start = c(1, 101, 201),
                        end = c(100, 200, 300),
                        globalState = c("M", "U", "M"),
                        u_eqFreq = c(0.1, 0.8, 0.1),
                        p_eqFreq = c(0.1, 0.1, 0.1),
                        m_eqFreq = c(0.8, 0.1, 0.8))
infoStr
```

## Parameter values

The default parameter values used by MethEvolSIM can be obtained using the function `get_model_parameters()`, that returns a data frame with the columns named as specified in the section "Model parameters":

```{r}
default_paramValues <- get_parameterValues()
default_paramValues
```

The section "Model parameters" provides further information about each of the parameters, including each parameter range of possible values. To simulate data with a different parameterization, the new value of a chosen parameter can be modified in the data frame.

```{r}
# Example: modification of parameter iota to value 0.2
default_paramValues$iota <- 0.2
default_paramValues
```

## Initial DNA Methylation Data

### Simulation

The initial data is generated in an object of class combiStructureGenerator.

```{r}
# Example with customized initial methylation frequencies and customized parameter values
custom_infoStr <- data.frame(start = c(1, 101, 201),
                        end = c(100, 200, 300),
                        globalState = c("M", "U", "M"),
                        u_eqFreq = c(0.1, 0.8, 0.1),
                        p_eqFreq = c(0.1, 0.1, 0.1),
                        m_eqFreq = c(0.8, 0.1, 0.8))
custom_params <- get_parameterValues()
custom_params$mu <- 0.005
initialD <- simulate_initialData(infoStr = custom_infoStr, params = custom_params)
# Returns customized parameters and simulated data
initialD$params
initialD$data
```

```{r}
# Example with simulated initial methylation frequencies and default parameter values
custom_infoStr <- data.frame(start = c(1, 101, 201),
                        end = c(100, 200, 300),
                        globalState = c("M", "U", "M"))
initialD <- simulate_initialData(infoStr = custom_infoStr)
# Returns default parameters
initialD$params
initialD$data
```

To get the parameter values of a combiStructureGenerator object, it can be given as argument to the function `get_parameterValues()`

```{r}
combiStr_object <- initialD$data
get_parameterValues(combiStr_object)
```

### Exploration of initial data

The methods of the class combiStructureGenerator can be used to access each of the contained structures:

```{r}
# E.g. access fist contained structure
combiStr_object$get_singleStr(1)
```

Each contained structure is an object of singleStructureGenerator and its methods can be used to get the contained information.

```{r}
# E.g. get the equilibrium frequencies of the first singleStructureGenerator
combiStr_object$get_singleStr(1)$get_eqFreqs()
```

```{r}
# E.g. get the sequence of methylation states of the second singleStructureGenerator
combiStr_object$get_singleStr(1)$get_seq()
```

## Simulate DNA Methylation Data Evolution

The evolution data is generated in an object of class treeMultiRegionSimulator.

```{r}
# Example with customized initial methylation frequencies, customized parameter values and default dt (0.01)
tree <- "((a:1, b:1):2, c:2, (d:3.7, (e:4, f:1):3):5);"
custom_infoStr <- data.frame(start = c(1, 101, 201),
                        end = c(100, 200, 300),
                        globalState = c("M", "U", "M"),
                        u_eqFreq = c(0.1, 0.8, 0.1),
                        p_eqFreq = c(0.1, 0.1, 0.1),
                        m_eqFreq = c(0.8, 0.1, 0.8))
custom_params <- get_parameterValues()
custom_params$mu <- 0.005
evolD <- simulate_evolData(infoStr = custom_infoStr, tree = tree, params = custom_params)
# Returns customized parameters, tree used, time step length for SSE process used (dt) and simulated data
evolD$data
evolD$dt
evolD$tree
evolD$params
```

If the evolution of given data at the root is simulated, and it wants to be done with customized parameter values, then the customized parameters are given to `simulate_initialData()` instead to `simulate_evolData()`.

```{r}
# Example with customized initial methylation frequencies, customized parameter values and default dt (0.01)
tree <- "((a:1, b:1):2, c:2, (d:3.7, (e:4, f:1):3):5);"
custom_infoStr <- data.frame(start = c(1, 101, 201),
                        end = c(100, 200, 300),
                        globalState = c("M", "U", "M"),
                        u_eqFreq = c(0.1, 0.8, 0.1),
                        p_eqFreq = c(0.1, 0.1, 0.1),
                        m_eqFreq = c(0.8, 0.1, 0.8))
custom_params <- get_parameterValues()
custom_params$mu <- 0.005
initialD <- simulate_initialData(infoStr = custom_infoStr, params = custom_params)$data
evolD <- simulate_evolData(rootData =initialD, tree = tree)
```
### Exploration of evolved data

The treeMultiStructureGenerator object contains a list `$Branch` with the combiStructureGenerator representing the data at the each node after a tree branch (but for the data at the root).

To know which offspring branches a node has, the combiStructureGenerator object method `$get_offspring_index()` can be used.

```{r}
treeData <- evolD$data
# E.g. offspring of data at root 
treeData$Branch[[1]]$get_offspring_index()
```
If the node is an inner node, its name is set as NULL. It can be accessed with the combiStructureGenerator method `$get_name()`

```{r}
treeData$Branch[[1]]$get_name()
```
But the leafs are named as in the given tree.

```{r}
treeData$Branch[[4]]$get_name()
```
If there were IWE events sampled in a branch, it can be accessed with the combiStructureGenerator method `$get_IWE_events()`

```{r}
treeData$Branch[[4]]$get_IWE_events()
```
In the case we want to know the equilibrium frequencies of a singleStructureGenerator object contained in the combiStructureGenerator we can use the combiStructureGenerator method to access the singleStructureGenerator object of interest `get_singleStr(i)` followed by the singleStructureGenerator method `get_eqFreqs()`

```{r}
# E.g. get the ewuilibrium frequencies of the second structure at the leaf "b"
treeData$Branch[[4]]$get_singleStr(2)$get_eqFreqs()
```

# Model parameters

## Initial methylation state distribution

Structures are categorized as CpG islands or non-islands. In the root of the genealogy we assume that each structure samples an equilibrium probability triple $(\pi_u,\pi_p, \pi_m)$. Current parameter values favor prevalent homogeneity in methylation levels within structures, most CpG islands with a majority predominantly unmethylated and a small proportion predominantly methylated, and most non-islands with a majority predominantly methylated and a small proportion predominantly unmethylated. The sampling of the equilibrium probability tripple is done using two distinct Beta distributions to sample the corresponding methylation frequencies.

**Island parameterization: alpha_pI, beta_pI, alpha_mI, beta_mI**

MethEvolSIM samples $\pi_p$ from the first Beta distribution. Current parameterization uses as initial parameter values a relatively small $\alpha_{I_p}$ and relatively large $\beta_{I_p}$ to encourage minor assignment of partially methylated states.

The second Beta distribution, scaled by $1 - \pi_p$,  corresponds to the proportion of methylated and unmethylated states. Currently both parameter values ($\alpha_{I_m}$ and $\beta_{I_m}$) set to small values while ${\alpha_{I_m} < \beta_{I_m}}$ to favour shifting the stochastic choice toward homogeneous states with a higher proportion of predominantly unmethylated islands, and assigned to $\pi_m$. Finally, we set $\pi_u$ as $1 - \pi_p - \pi_m$.

**Non-island parameterization: alpha_pNI, beta_pNI, alpha_mNI, beta_mNI**

Following the same rationale as for island parameterization, all the regions not belonging to an island sample a common equilibrium probability triple from two Beta distributions with parameters $\alpha_{NI_p}$, $\beta_{NI_p}$, $\alpha_{NI_m}$ and $\beta_{NI_m}$. The parameter values are currently set to favor the characteristic high overall methylation level.

You can simulate DNA methylation along a tree given some data at root, or by simulating data at root given a structural distribution and letting it evolve.

**Parameter ranges** 
The alpha and beta parameters of a Beta distribution must be greater than 0, that is (0, $\infty$).

## DNA methylation evolution processes

DNA methylation state evolution happens by two structure-specific events: CpG single site events (SSEs) and CpG island wide events (IWEs).

### Island-wide events (IWEs)

IWEs change the island methylation probabilities and some of the CpG sites in the CpG island can simultaneously change their state at the same time. Different CpG islands, however, are assumed to evolve independently of each other. IWEs occur per CpG island at a rate denoted by $\mu$. Consequently, over a branch of length $l$, the expected number of IWEs per island is given by $\mu \cdot l$.

**IWE parameterization: mu. Parameter range:**
Should be a small number, but it can take any value from 0, that is [0, $\infty$)

### Single-site events (SSEs)

In addition to IWEs we allow for single site events (SSEs), which change the methylation states of single CpG sites within and between CpG islands. In our model SSEs account for two dynamics: independent of each other (SSEi), and  assuming correlations between adjacent sites (SSEc).

As SSEc transition probabilities are dependent on adjacent sites we need to approximate SSE based evolution by short time steps, allowing for CpG sites to actualize their state. Therefore, we first discretize the branch intervals between IWEs in time intervals of length 0.01. For each CpG position and each time interval the rate matrix for the transitions between the states unmethylated, partially methylated and methylated results from the addition of the rate matrices of the two types of SSE events, so that $\mathbb{E}(R_i + R_c)=1$, $R_i$ and $R_c$ representing the rate factors of independent and correlated processes. The proportion of each of the two SSE types is determined by the parameter $\iota$.

Different positions $j$ have different SSEi rates $R_{i,j}$ coming from a discretized gamma distribution with 3 categories, expected value $\iota$ and shape parameters $\alpha_{R_{i}}$ and ${\beta_{R_{i}}}=\frac{\alpha_{R_{i}}}{\iota}$. The probability to be in each respective rate category is 1/3. For the rate of collaborative SSEs $R_c = 1-\iota$, the probability of considering left of right neighbour is equal.

**SSE process parameterization: iota, alpha_Ri**
**Parameter ranges** 
The alpha parameter of a Gamma distribution must be greater than 0, that is (0, $\infty$). Iota must be between 0 (non-included) and 1, that is (0,1].