### 2024-11-25 extensive_tests/draft-abc-inferences-vignette.Rmd
Adding draft of vignettes and snakemake workflow

---
title: "ABC inferences"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Introduction to MethEvolSIM}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, echo=FALSE}
## TODO: Update 
library(devtools)
load_all()
```

# Worflow for simulation-based inferences

This workflow uses [Snakemake](https://snakemake.readthedocs.io) to manage the steps for simulating DNA methylation patterns, assessing ABC accuracy, and inferring model parameters.

## Requirements

- Snakemake installed (`conda create -c conda-forge -c bioconda -n snakemake snakemake`)

- R packages: optparse (add to dependencies), parallel

- Access to the provided `Snakefile` and associated scripts.

- An .RData file with a dataframe specifying the distribution of methylation sites in the genomic region.

## Steps in the Workflow

1. Save as simulation design the genomic structure, tree(s) -- GIVEN -- and sampled parameter combinations.

2. Run the simulations. TODO: Before submitting load package without devtools

3. Compute the selected summary statistics. TODO: When added sumStats functions to package, stop sourcing script.

4. Assess ABC accuracy using cross-validation.

5. Infer parameters from empirical DNA methylation data.