### 2024-12-05 extensive_tests/draft-sumstat-vignette.Rmd
Updating changes in sumstats

---
title: "Summary Statistics"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Introduction to MethEvolSIM}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, echo=FALSE}
## TODO: Update 
library(devtools)
load_all()
```

Interpretation of arguments:

 - tip can be interpreted as sample or replicate

 - structure can be interpreted as island or non-island

## Frequency of methylation states per genomic structure type

### Global Mean

- **Partially-methylated sites**: Let $x_{i,t}$ and $y_{j,t}$ be the frequency of methylation state $p$ in island $i \in\{1, \cdots, I\}$ and non-island $j \in\{1, \cdots, J\}$ at tip $t$, respectively. We define the mean global frequency of $x$ and $y$ as:

\[
\overline{x}=\frac1I\sum_{i=1}^{I}\frac1T\sum_{t=1}^{T}x_{i,t} \\
\overline{y}=\frac1J\sum_{j=1}^{J}\frac1T\sum_{t=1}^{T}y_{j,t}
\]

, where $x_{i,t}$ is the frequency of state $p$ in island $i$ and tip $t$ and $y_{j,t}$ is the frequency of state $p$ in non-island $j$ and tip $t$.

- **Methylated sites**: Let $w_{i,t}$ and $z_{j,t}$ be the frequency of methylation state $m$ in island $i \in\{1, \cdots, I\}$ and non-island $j \in\{1, \cdots, J\}$ at tip $t$, respectively. We define the mean global frequency of $w$ and $z$ as:

\[
\overline{w}=\frac1I\sum_{i=1}^{I}\frac1T\sum_{t=1}^{T}w_{i,t} \\
\overline{z}=\frac1J\sum_{j=1}^{J}\frac1T\sum_{t=1}^{T}z_{j,t}
\]

, where $w_{i,t}$ is the frequency of state $p$ in island $i$ and tip $t$ and $z_{j,t}$ is the frequency of state $m$ in non-island $j$ and tip $t$.

These values can be computed with the functions `get_islandMeanFreqP`, `get_nonislandMeanFreqP`,`get_islandMeanFreqM` and `get_nonislandMeanFreqM` as in the following examples:

```{r}
# 1 tip / sample / replicate
sample_n <- 1
index_islands <- c(1, 3)
index_nonislands <- c(2, 4)
data <- list(c(.5, .5, 0, 0, 0, .5), c(.5, 0, 0, .5), c(.5, .5, 0), c(0, 0, .5)) # tip 1
get_islandMeanFreqP(index_islands, data, sample_n)
get_nonislandMeanFreqP(index_nonislands, data, sample_n)
get_islandMeanFreqM(index_islands, data, sample_n)
get_nonislandMeanFreqM(index_nonislands, data, sample_n)
```

```{r}
# 2 tip / sample / replicate
sample_n <- 2
index_islands <- c(1, 3)
index_nonislands <- c(2, 4)
data <- list(
  list(c(.5, .5, 0, 0, 0, .5), c(.5, 0, 0, .5), c(.5, .5, 0), c(0, 0, .5)), # tip 1
  list(c(0, .5, .5, 1, 1, .5), c(1, .5, 1, .5), c(0, .5, .5), c(1, .5, 1)) # tip 2
  )
get_islandMeanFreqP(index_islands, data, sample_n)
get_nonislandMeanFreqP(index_nonislands, data, sample_n)
get_islandMeanFreqM(index_islands, data, sample_n)
get_nonislandMeanFreqM(index_nonislands, data, sample_n)
```

### Mean SD of observed frequencies per structure type at tree tips

- **Partially-methylated sites**: Let $x_{i,t}$ and $y_{j,t}$ be the frequency of methylation state $p$ in island $i \in\{1, \cdots, I\}$ and non-island $j \in\{1, \cdots, J\}$, respectively. We define the standard deviation of the frequency of $x$ and $y$ at tip $t$ as:

\[
\sigma_{t}(x)=\sqrt{\frac{1}{I-1}\sum_{i=1}^{I}(x_{i,t}-\overline{x_{.,t}})^2} \\
\sigma_{t}(y)=\sqrt{\frac{1}{J-1}\sum_{j=1}^{J}(y_{i,t}-\overline{y_{.,t}})^2}
\]

, where $\overline{x_{.,t}}$ is the mean frequency of state $p$ at tip $t$ in islands and $\overline{y_{.,t}}$ is the mean frequency of state $p$ at tip $t$ in non-islands:

\[
\overline{x_{.,t}}=\frac{1}{I}\sum_{i=1}^{I}x_{i,t} \\
\overline{y_{.,t}}=\frac{1}{J}\sum_{j=1}^{J}x_{j,t}
\]

Therefore, the mean standard deviation across tips is:

\[
\hat{\sigma}(x)=\frac{1}{T}\sum_{t=1}^{T}\sigma_t(x) \\
\hat{\sigma}(y)=\frac{1}{T}\sum_{t=1}^{T}\sigma_t(y) \\
\]

- **Methylated sites**: Let $w_{i,t}$ and $z_{j,t}$ be the frequency of methylation state $m$ in island $i \in\{1, \cdots, I\}$ and non-island $j \in\{1, \cdots, J\}$, respectively. We define the standard deviation of the frequency of $w$ and $z$ at tip $t$ as:

\[
\sigma_{t}(w)=\sqrt{\frac{1}{I-1}\sum_{i=1}^{I}(x_{i,t}-\overline{x_{.,t}})^2} \\
\sigma_{t}(z)=\sqrt{\frac{1}{J-1}\sum_{j=1}^{J}(y_{i,t}-\overline{y_{.,t}})^2}
\]

, where $\overline{x_{.,t}}$ is the mean frequency of state $m$ at tip $t$ in islands and $\overline{z_{.,t}}$ is the mean frequency of state $m$ at tip $t$ in non-islands:

\[
\overline{w_{.,t}}=\frac{1}{I}\sum_{i=1}^{I}w_{i,t} \\
\overline{z_{.,t}}=\frac{1}{J}\sum_{j=1}^{J}z_{j,t}
\]

Therefore, the mean standard deviation across tips is:

\[
\hat{\sigma}(w)=\frac{1}{T}\sum_{t=1}^{T}\sigma_t(w) \\
\hat{\sigma}(z)=\frac{1}{T}\sum_{t=1}^{T}\sigma_t(z) \\
\]

```{r}
# 1 tip / sample / replicate
sample_n <- 1
index_islands <- c(1, 3)
index_nonislands <- c(2, 4)
data <- list(c(.5, .5, 0, 1, 1, .5), c(.5, 0, 1, .5), c(.5, .5, 0), c(0, 0, .5)) # tip 1
get_islandSDFreqP(index_islands, data, sample_n)
get_nonislandSDFreqP(index_nonislands, data, sample_n)
get_islandSDFreqM(index_islands, data, sample_n)
get_nonislandSDFreqM(index_nonislands, data, sample_n)
```

```{r}
# 2 tip / sample / replicate
sample_n <- 2
index_islands <- c(1, 3)
index_nonislands <- c(2, 4)
data <- list(
  list(c(.5, .5, 0, 0, 0, 1), c(.5, 0, 0, .5), c(1, .5, 0), c(0, 0, .5)), # tip 1
  list(c(0, .5, .5, 1, 1, .5), c(1, .5, 1, .5), c(0, .5, .5), c(1, .5, 1)) # tip 2
  )
get_islandSDFreqP(index_islands, data, sample_n)
get_nonislandSDFreqP(index_nonislands, data, sample_n)
get_islandSDFreqM(index_islands, data, sample_n)
get_nonislandSDFreqM(index_nonislands, data, sample_n)
```

Note that we average across tips as there can be correlations between the mean standard deviation of tips (e.g. under the case of an event affecting the states at two tips).

# Mean correlation of a segment of methylation states per genomic structure type

Let $s$ be the length of the 2 adjacent segments of central positions to consider.
Let $l$ represent a number of positions to exclude at the start and end of each island/non-island structure.
Define the index positions $p=l+1$ and $p' = s-l$.
Let $x_{i,t,[p,p'-1]}$ and $x_{i,t,[p+1,p']}$ be the methylation states $x \in \{u, p, m\}$ in island $i \in \{1, \cdots ,I\}$ at tip $t$ for the given segments. 
Let $y_{j,t,[p,p'-1]}$ and $y_{j,t,[p+1,p']}$ be the methylation states $y \in \{u, p, m\}$ in non-island $j \in \{1, \cdots ,J\}$ at tip $t$ for the given segments.

We define the mean correlation as:

\[
\overline{Cor(x)_{s,l}}=\frac1I\frac1T\sum_{i=1}^{I}\sum_{t=1}^{T}Cor(x_{i,t,[p:p'-1]}, x_{i,t,[p+1:p']}) \\
\overline{Cor(y)_{s,l}}=\frac1J\frac1T\sum_{j=1}^{J}\sum_{t=1}^{T}Cor(y_{j,t,[p:p'-1]}, y_{j,t,[p+1:p']})
\]

where correlations are computed only for segment pairs where the standard deviation of each segment is non-zero:

\[
\sigma(x_{i,t,[p:p'-1]}) \neq 0 \text{ and } \sigma(x_{i,t,[p+1:p']})\neq 0 \\
\sigma(y_{j,t,[p:p'-1]}) \neq 0 \text{ and } \sigma(y_{j,t,[p+1:p']})\neq 0
\]

In this formulation:

- Only structures with a minimum length of $n = s + 2 \cdot l$ are considered.

- If $l = 0$, all the positions are considered.

```{r}
# 1 tip / sample / replicate
sample_n <- 1
index_islands <- c(1, 3)
index_nonislands <- c(2, 4) 
data <- list(c(.5, 0, 0, 0, .5, .5, .5, .5, .5, 1, .5, 0, 0, 0, .5, .5, .5, .5, .5, 1, .5, 0, 0, 0, .5, .5, .5, .5, .5, 1), # 30 sites
               c(.5, 1, 1, 1, .5, .5, 1, 1, 1, .5, .5, 1, 1, 1, .5, .5, 1, 1, 1, .5, .5, 1, 1, 1, .5), # 25 sites
               c(.5, 0, 0, .5, 1, .5, 0, 0, .5, 1, .5, 0, 0, .5, 1, .5, 0, 0, .5, 1, .5, 0, 0, .5, 1, .5, 0, 0, .5, 1, .5, 0, 0, .5, 1, .5, 0, 0, .5, 1), # 40 sites
               c(1, 1, 1, .5, .5, .5, 0, 0, 0, .5, 1, 1, 1, .5, .5, .5, 0, 0, 0, .5, 1, 1, 1, .5, .5, .5, 0, 0, 0, .5, .5, 0, 0, 0, .5)) # 35 sites
compute_meanCor_i(index_islands, minN_CpG = 10, shore_length = 5, data, sample_n = 1)
compute_meanCor_ni(index_nonislands, minN_CpG = 10, shore_length = 5, data, sample_n = 1)
```

```{r}
# 2 tip / sample / replicate
sample_n <- 2
index_islands <- c(1, 3)
index_nonislands <- c(2, 4)
data <- list(
  # tip 1
    list(c(.5, 0, 0, 0, .5, .5, .5, .5, .5, 1, .5, 0, 0, 0, .5, .5, .5, .5, .5, 1, .5, 0, 0, 0, .5, .5, .5, .5, .5, 1), # 30 sites
         c(.5, 1, 1, 1, .5, .5, 1, 1, 1, .5, .5, 1, 1, 1, .5, .5, 1, 1, 1, .5, .5, 1, 1, 1, .5), # 25 sites
         c(.5, 0, 0, .5, 1, .5, 0, 0, .5, 1, .5, 0, 0, .5, 1, .5, 0, 0, .5, 1, .5, 0, 0, .5, 1, .5, 0, 0, .5, 1, .5, 0, 0, .5, 1, .5, 0, 0, .5, 1), # 40 sites
         c(1, 1, 1, .5, .5, .5, 0, 0, 0, .5, 1, 1, 1, .5, .5, .5, 0, 0, 0, .5, 1, 1, 1, .5, .5, .5, 0, 0, 0, .5, .5, 0, 0, 0, .5)), # 35 sites
  # tip 2
    list(c(.5, 0, 0, .5, .5, .5, 0, 0, .5, 1, .5, 0, 0, 0, 0, .5, .5, 1, 1, 1, .5, 0, 0, 0, .5, .5, 1, 1, 1, 1), # 30 sites
         c(.5, .5, 1, 1, .5, .5, 1, 1, 1, .5, .5, 0, 0, 0, .5, .5, 1, 1, 1, .5, .5, 1, 1, 1, .5), # 25 sites
         c(.5, 0, 0, .5, 1, .5, 0, 0, .5, 1, .5, 0, 0, .5, .5, .5, 0, 0, .5, 1, 1, 1, 1, .5, 1, .5, 0, 0, .5, 1, .5, 0, 0, .5, 1, .5, 0, 0, .5, 1), # 40 sites
         c(1, 1, 1, .5, .5, .5, 0, 0, 0, .5, 1, 1, 1, 1, .5, .5, 0, 0, 0, .5, 1, 1, 1, .5, .5, .5, .5, .5, 0, .5, .5, .5, .5, 0, .5)) # 35 sites
  )
compute_meanCor_i(index_islands, minN_CpG = 10, shore_length = 5, data, sample_n = 2)
compute_meanCor_ni(index_nonislands, minN_CpG = 10, shore_length = 5, data, sample_n = 2)
```

*Correlation measures how much the state of each site says about the next. So, when there is no variation (e.g. in middle segment state u of a site is followed by state u of the next), there is no correlation.*

## Comparison of tree tips

### Frequency of Methylation State Changes at Single Sites

We compare the methylation state of single sites at the tree cherries. A cherry is a pair of leaf nodes (also called tips or terminal nodes) in a phylogenetic tree that share a direct common ancestor. In other words, if two leaves are connected to the same internal node and no other leaves are connected to that internal node, they form a cherry.

get_FChange_cherryData (uses find_cherries and cherry_diffs)