### 2024-12-09 extensive_tests/CFTP_convergence/draft-test_CFTP_convergence.Rmd
Adding files for test CFTP convergence workflow for reproducibility

---
title: "CFTP Convergence"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Test for CFTP Convergence}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, echo=FALSE}
## TODO: Update 
library(devtools)
load_all()
```

# Worflow for simulation-based inferences

This workflow uses [Snakemake](https://snakemake.readthedocs.io) to manage the steps for simulating DNA methylation patterns, assessing ABC accuracy, and inferring model parameters.

## Requirements

- Snakemake installed (`conda create -c conda-forge -c bioconda -n snakemake snakemake`)

- R packages:

- Access to the provided `Snakefile` and associated scripts.

- An .RData file with a dataframe specifying the distribution of methylation sites in the genomic region.

## Steps in the Workflow

1. Bla

### Scripts

- `run_sim.R`. TODO: Update way in which the package is loaded before submitting.