### 2024-12-16 extensive_tests/abc_inferences/draft-abc-inferences-vignette.Rmd
Updating script get_spatial_str of abc_inferences workflow

---
title: "ABC inferences"
output: 
  rmarkdown::html_vignette:
    toc: true
bibliography: references.bib
vignette: >
  %\VignetteIndexEntry{ABC Inferences with MethEvolSIM}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,  # Set the figure width to match the text width
  fig.height = 5, # Set the figure height
  out.width = "100%"  # Ensure it takes up 100% of the available width in HTML
)
```

```{r, echo=FALSE}
## TODO: Update 
library(devtools)
load_all()
```

# Workflow dependencies

Are listed under the file `environment.yaml`. Before running the workflow, a conda environment can be created with:

```{bash, eval = FALSE}
conda env create -n abc_inferences -f environment.yaml
```

# Selection of genomic region(s)

-Requires: CGI annotation file (`.txt` file) obtained with makeCGI v1.3.4 [@hiddenMarkov]

- The genome assembly file: Obtain the FASTA file ([GCF_000238955.4_M_zebra_UMD2a_genomic.fa](https://www.ncbi.nlm.nih.gov/datasets/genome/GCF_000238955.4/)) from NCBI (file source *RefSeq*, file type *Genome Sequences (FASTA)*)

Both under subdirectory genomic_dist

```{r}
# Load data
file_path <- "genomic_dist/CGI-Mzebra.txt" # Adjust the path if necessary
cgi_data <- read.table(file_path, header = TRUE, sep = "\t")

# View the first lines
head(cgi_data)
# View the last lines
tail(cgi_data)
```

The prefixes `NC_` and `NW_` in chromosome or genomic sequence identifiers come from the RefSeq database and indicate different types of assembly or annotation status for genomic sequences:

- `NC_` refers to non-redundant complete genomic molecules, which typically represent well-characterized, curated, and assembled sequences.

- `NW_` refers to scaffolds or unplaced sequences that are part of the genome assembly but cannot yet be confidently assigned to a specific chromosome or region.

```{r}
# Count the number of CGIs per chromosome
cgi_count <- table(cgi_data$chr)

# View results grouped by type of chromosome assembly

# Split into groups based on chromosome prefix
groups <- ifelse(grepl("^NC_", names(cgi_count)), "NC", "NW")

# Create a dataframe
cgi_data <- data.frame(
  group = groups,
  count = as.numeric(cgi_count)
)

# Plot boxplots
boxplot(count ~ group, data = cgi_data,
        col = c("skyblue", "salmon"),
        main = "Distribution of Counts for NC_ and NW_ Chromosomes",
        xlab = "Chromosome Type",
        ylab = "Counts")

```

## Fixed DMR found in paper: Associated genes

Add reference

- *Smad4a*: LOC101481185. Fixed but not associated with differential expression.

Genes reported as fixed DMR with significant transcriptional downregulation in benthic livers and associated with fixed benthic-specific hypermethylated levels at their promoters, suggesting retained epigenetic divergence associated with different diets.

- *Dpp4*: Part of the insulin methabolic pathway. Location according to NCBI annotation release 104 (LOC101483164): NC_036801.1 (29502362..29517382)

- *Ces2*: Liver carboxylesterase 2, part of the fatty acid methabolic pathway. Location according to NCBI annotation release 104 (LOC101473569):  	NW_020192811.1 (49164..53741, complement)

Lets explore *Dpp4*, as it is associated with differential gene expression and it is located in a better characterized genomic region:

```{r}
dpp4_chr <- "NC_036801.1"
dpp4_start <- 29502362
dpp4_end <- 29517382
```

## Explore the CpG counts in islands and non islands generated from the reference genome and the CGI annotation

```{r}
load("genomic_dist/CpG_count.RData")
```

```{r}
# Filter the data for the given chromosome and region
filtered_data <- CpG_count[CpG_count$chr == dpp4_chr & CpG_count$start >= dpp4_start - 5e5 & CpG_count$start <= dpp4_end + 5e5, ]

# Set up the plot area
plot(filtered_data$start, filtered_data$CpG_count, 
     type = "n", # No points, just the axis
     xlab = "Position (bp)", ylab = "CpG Count", 
     main = "CpG Count by Island/Non-Island Regions", 
     xlim = c(min(filtered_data$start), max(filtered_data$end)), 
     ylim = c(0, max(filtered_data$CpG_count) * 1.1))

# Add bars with color based on 'str' (island vs non-island)
for (i in 1:nrow(filtered_data)) {
  # Calculate the width of each region (difference between start and end)
  region_width <- filtered_data$end[i] - filtered_data$start[i]

  if (filtered_data$str[i] == "Non-island") {
    # Non-island (blue)
    rect(filtered_data$start[i], 0, filtered_data$end[i], 
         filtered_data$CpG_count[i], col = "blue", border = "blue")
  } else {
    # Island (red)
    rect(filtered_data$start[i], 0, filtered_data$end[i], 
         filtered_data$CpG_count[i], col = "red", border = "red")
  }
}

# Add vertical lines showing where the gene is
abline(v = dpp4_start)
abline(v = dpp4_end)

# Add a legend
legend("topright", legend = c("Non-Island", "Island"), 
       fill = c("blue", "red"), title = "Island Type")

```

Genomic region close to gene *dpp4*. X-axis corresponds to the positions in chromosome `r dpp4_chr`. Y-axis corresponds to CpG counts at each genomic region. Each genomic region is represented by a bar starting and ending in the corresponding chromosome position. Vertical black lines represent start and end of gene *dpp4*.

```{r}

# Filter the data for the given chromosome and region
filtered_data <- CpG_count[CpG_count$chr == dpp4_chr & CpG_count$start > dpp4_start - 5e5 & CpG_count$end < dpp4_end + 5e5, ]

# Calculate the CpG density for each region (CpG count / region length)
filtered_data$CpG_density <- filtered_data$CpG_count / (filtered_data$end - filtered_data$start)

# Plot CpG density vs. position (using 'start' as x-axis)

# Set up the plot area with no points initially
plot(filtered_data$start, filtered_data$CpG_density, 
     type = "n", # No points, just the axis
     xlab = "Position (bp)", ylab = "CpG Density", 
     main = "CpG Density by Island/Non-Island Regions", 
     xlim = c(min(filtered_data$start), max(filtered_data$end)), 
     ylim = c(0, max(filtered_data$CpG_density) * 1.1))

# Add points colored by 'str' (0 for non-island, 1 for island)
points(filtered_data$start, filtered_data$CpG_density, 
       col = ifelse(filtered_data$str == 1, "red", "blue"), 
       pch = 16)  # pch=16 gives filled circles

# Add vertical lines showing where the gene is
abline(v = dpp4_start)
abline(v = dpp4_end)

# Add a legend
legend("topright", legend = c("Non-Island", "Island"), 
       fill = c("blue", "red"), title = "Type")

```

Genomic region close to gene *dpp4*. X-axis corresponds to the positions in chromosome `r dpp4_chr`. Y-axis corresponds to CpG density at each genomic region. Each genomic region is represented by a dot at its corresponding genome start position. Vertical black lines represent start and end of gene *dpp4*.

When using this region the number of CpG sites to simulate would be:

```{r}
sum(filtered_data$CpG_count)
```

with number of non-island and island structures:

```{r}
# Count the number of occurrences of each factor level in the 'str' column
table(filtered_data$str)
```

# Worflow for simulation-based inferences

This workflow uses [Snakemake](https://snakemake.readthedocs.io) to manage the steps for simulating DNA methylation patterns, assessing ABC accuracy, and inferring model parameters.

## Requirements

- Snakemake installed (`conda create -c conda-forge -c bioconda -n snakemake snakemake`)

- R packages: optparse (add to dependencies), parallel, ape

- Access to the provided `Snakefile` and associated scripts.

- An .RData file with a dataframe specifying the distribution of methylation sites in the genomic region.

- An .RData file with a tree in ape format (TODO: modify this and check whether scrm is also producing trees in the same format)

## Steps in the Workflow

1. Save as simulation design the genomic structure, tree(s) -- GIVEN -- and sampled parameter combinations.

2. Run the simulations. TODO: Before submitting load package without devtools

3. Compute the selected summary statistics. TODO: When added sumStats functions to package, stop sourcing script.

4. Assess ABC accuracy using cross-validation.

5. Infer parameters from empirical DNA methylation data.

# References