### 2024-12-18 extensive_tests/CFTP_convergence/draft-test_CFTP_convergence.Rmd
Updating snakemake workflow for test CFTP convergence until run_sim

---
title: "CFTP Convergence"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Test for CFTP Convergence}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, echo=FALSE}
## TODO: Update 
library(devtools)
load_all()
```

# Worflow for simulation-based inferences

This workflow uses [Snakemake](https://snakemake.readthedocs.io) to manage the steps for simulating DNA methylation patterns, assessing ABC accuracy, and inferring model parameters.

## Requirements

- Snakemake installed (`conda create -c conda-forge -c bioconda -n snakemake snakemake`)

- Dependencies are on: `environment.yaml`

- Access to the provided `Snakefile` and associated scripts.

- An .RData file with a dataframe specifying the distribution of methylation sites in the genomic region (under the same Snakemake directory)

-

## Steps in the Workflow

1. Bla

### Scripts

- `run_sim.R`. TODO: Update way in which the package is loaded before submitting. Uses as many cores as the config param n_sim

## Running the workflow with nohup

When using nohup one needs to activate the environment in the command.

```{bash, eval = FALSE}
# E.g. running the run_sim rule
nohup conda run -n cftp_testConvergence snakemake run_sim &
```