### 2024-12-24 application_examples/abc_inferences/archipelago-definition.Rmd
Adding file archipelago-definition

---
title: "Definition of Archipelagos"
author: "Sara"
date: "`r Sys.Date()`"
output: html_document
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# If missing, install required packages

if (!requireNamespace("BiocManager", quietly = TRUE)) {
  install.packages("BiocManager")
}

if (!requireNamespace("rtracklayer", quietly = TRUE)) {
  BiocManager::install("rtracklayer")
}

library(rtracklayer)
```

# Archipelago definition

We define archipelagos as genomic regions containing clusters of CGIs. We can interpret archipelagos as regions of the genome larger than CGI that together may regulate a gene or set of genes. Therefore, an approach to define archipelagos can be to choose thresholds that lead to clustering CGI in a way so that each archipelago overlaps with a promoter region (or maximizes cases of overlap single-archipelago and promoter).

Our approach then has the following steps:

1. Definition of promoter regions.
2. Filtering of CGI by promoter overlap.

## Definition of promoter regions

As in [@vernaz2022epigenetic], since no functional annotation exists for Lake Malawi cichlid genomes, we define promoter regions in silico as regions ±1 kbp around the TSS.

For that, we use [Annotation Features File (GTF) from the NCBI RefSeq Assembly](https://www.ncbi.nlm.nih.gov/datasets/genome/GCF_000238955.4/)

```{r}
# Load the GTF file
gtf_file <- "/home/sara/methylation/MethEvolSIM/application_examples/abc_inferences/genomic_dist/ncbi_dataset/ncbi_dataset/data/GCF_000238955.4/genomic.gtf" 
gtf <- import(gtf_file)
head(gtf)
```

The gtf file is structured as a GRanges object containing genomic annotation information with the following key components (columns):

- `seqnames`: Chromosome or scaffold name (e.g., NC_036780.1).

- `ranges`: Start and end positions of the genomic feature.

- `strand`: Strand of the feature (+ for forward strand, - for reverse strand).

- `source`: Annotation source (e.g., Gnomon).

- `type`: Feature type (e.g., gene, transcript, exon).

- `Metadata`: Additional columns include gene_id, transcript_id, gene, product, gene_biotype, exon_number, etc., which describe the feature in detail.

Example Entries:

- Row 1: A gene on the + strand, identified as LOC106676409, classified as lncRNA.

- Row 2: A transcript corresponding to the same gene.

- Rows 3-5: Exons belonging to the transcript.

- Row 6: A gene on the - strand, classified as protein_coding.

We filter for transcripts:

```{r}
# Filter for transcripts
gtf_transcripts <- gtf[gtf$type == "transcript", ]
head(gtf_transcripts)
```

And define the TSS by accounting for the strand orientation so that:

- For Features on the + Strand: The TSS is the smallest genomic coordinate (start).

- For Features on the - Strand: The TSS is the largest genomic coordinate (end), as transcription occurs in the reverse direction.

```{r}
# Extract TSS (assuming + strand)
promoters <- data.frame(
  chr = seqnames(gtf_transcripts),
  tss = ifelse(strand(gtf_transcripts) == "+", start(gtf_transcripts), end(gtf_transcripts)),
  strand = strand(gtf_transcripts)
)
```