### 2025-01-07 extensive_tests/CFTP_convergence/draft-test_CFTP_convergence.Rmd
Updating draft of vignette CFTP convergence

---
title: "CFTP Convergence"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Test for CFTP Convergence}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# CFTP

Theoretical intro.

## Implementation

Currently neglects IWEs but allows to sample a sequence of methylation states from the equilibrium given by the equilibrium frequencies of the genomic regions, the correlated and independent SSE.

## Test

This workflow uses [Snakemake](https://snakemake.readthedocs.io) to manage the steps for simulating DNA methylation patterns, assessing ABC accuracy, and inferring model parameters.

**Add summary**

### Requirements

-   Dependencies are listed on the file `environment.yaml`. Before running the workflow, a conda environment can be created with:

```{bash, eval = FALSE}
conda env create -n cftp_testConvergence -f environment.yaml
```

Then, the workflow can be run by:

```{bash, eval = FALSE}
conda activate cftp_testConvergence
snakemake --use-conda
```

Or, if using `nohup` by:

```{bash, eval = FALSE}
conda activate cftp_testConvergence
nohup nice snakemake --use-conda run_sim > run_sim.out &
```

-   Provided `Snakefile`, `config.yaml` and associated scripts.

## Steps in the Workflow

All the given parameters to the workflow steps are taken from the `config.yaml` file.

1.  Rule `get_spatial_str` run by script `get_spatial_str.R`. The test simulates data on a simplified genomic structure alternating island and non-island regions each a given number of times and each (`Str_n`) with a constant number of CpGs  (`CpG_n`). The output is a file named as `methSite_genomicDist` under the directory `dir`.

2.  Rule `set_design` run by script `set_sim_design.R`. Takes as input the output of the previous rule. For a number of model parameter combinations `n_sim` and setting a seed `seed` for reproducibility, samples the parameter combinations from the prior distributions as defined in `prior_distributions`. The output is a file named "design-RData" under the directory `dir` containing both the spatial structure defined in the first rule and the sampled parameters. *Note that this script already corrects the values of* $\iota$ *and* $\alpha_{R_i}$ *in the sampled parameters to be a minimum of 1e-2*.

### Scripts

-   `run_sim.R`. TODO: Update way in which the package is loaded before submitting. Uses as many cores as the config param n_sim

## Running the workflow with nohup

When using nohup one needs to activate the environment in the command.

```{bash, eval = FALSE}
# E.g. running the run_sim rule
nohup conda run -n cftp_testConvergence snakemake run_sim &
```

## Results

```{r, fig.align='center',  out.width='50%'}
knitr::include_graphics("Figures/CFTP_testConvergence_paramsID_01.pdf")
```

```{r, fig.align='center',  out.width='50%'}
knitr::include_graphics("Figures/CFTP_testConvergence_paramsID_02.pdf")
```

```{r, fig.align='center',  out.width='50%'}
knitr::include_graphics("Figures/CFTP_testConvergence_paramsID_03.pdf")
```

```{r, fig.align='center',  out.width='50%'}
knitr::include_graphics("Figures/CFTP_testConvergence_paramsID_04.pdf")
```

```{r, fig.align='center',  out.width='50%'}
knitr::include_graphics("Figures/CFTP_testConvergence_paramsID_05.pdf")
```

```{r, fig.align='center',  out.width='50%'}
knitr::include_graphics("Figures/CFTP_testConvergence_paramsID_06.pdf")
```

```{r, fig.align='center',  out.width='50%'}
knitr::include_graphics("Figures/CFTP_testConvergence_paramsID_07.pdf")
```

```{r, fig.align='center',  out.width='50%'}
knitr::include_graphics("Figures/CFTP_testConvergence_paramsID_08.pdf")
```

```{r, fig.align='center',  out.width='50%'}
knitr::include_graphics("Figures/CFTP_testConvergence_paramsID_09.pdf")
```

```{r, fig.align='center',  out.width='50%'}
knitr::include_graphics("Figures/CFTP_testConvergence_paramsID_10.pdf")
```