### 2025-03-01 extensive_tests/draft-sumstat-vignette.Rmd
Updating functions for fitch estimation per island and corresponding tests

---
title: "Summary Statistics"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Summary Statistics with MethEvolSIM}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, echo=FALSE}
## TODO: Update 
library(devtools)
load_all()
```

Interpretation of arguments:

 - tip can be interpreted as sample or replicate

 - structure can be interpreted as island or non-island

## Frequency of methylation states per genomic structure type

### Global Mean

- **Partially-methylated sites**: Let $x_{i,t}$ and $y_{j,t}$ be the frequency of methylation state $p$ in island $i \in\{1, \cdots, I\}$ and non-island $j \in\{1, \cdots, J\}$ at tip $t$, respectively. We define the mean global frequency of $x$ and $y$ as:

\[
\overline{x}=\frac1I\sum_{i=1}^{I}\frac1T\sum_{t=1}^{T}x_{i,t} \\
\overline{y}=\frac1J\sum_{j=1}^{J}\frac1T\sum_{t=1}^{T}y_{j,t}
\]

, where $x_{i,t}$ is the frequency of state $p$ in island $i$ and tip $t$ and $y_{j,t}$ is the frequency of state $p$ in non-island $j$ and tip $t$.

- **Methylated sites**: Let $w_{i,t}$ and $z_{j,t}$ be the frequency of methylation state $m$ in island $i \in\{1, \cdots, I\}$ and non-island $j \in\{1, \cdots, J\}$ at tip $t$, respectively. We define the mean global frequency of $w$ and $z$ as:

\[
\overline{w}=\frac1I\sum_{i=1}^{I}\frac1T\sum_{t=1}^{T}w_{i,t} \\
\overline{z}=\frac1J\sum_{j=1}^{J}\frac1T\sum_{t=1}^{T}z_{j,t}
\]

, where $w_{i,t}$ is the frequency of state $p$ in island $i$ and tip $t$ and $z_{j,t}$ is the frequency of state $m$ in non-island $j$ and tip $t$.

These values can be computed with the functions `get_islandMeanFreqP`, `get_nonislandMeanFreqP`,`get_islandMeanFreqM` and `get_nonislandMeanFreqM` as in the following examples:

```{r}
# 1 tip / sample / replicate
sample_n <- 1
index_islands <- c(1, 3)
index_nonislands <- c(2, 4)
data <- list(c(.5, .5, 0, 0, 0, .5), c(.5, 0, 0, .5), c(.5, .5, 0), c(0, 0, .5)) # tip 1
get_islandMeanFreqP(index_islands, data, sample_n)
get_nonislandMeanFreqP(index_nonislands, data, sample_n)
get_islandMeanFreqM(index_islands, data, sample_n)
get_nonislandMeanFreqM(index_nonislands, data, sample_n)
```

```{r}
# 2 tip / sample / replicate
sample_n <- 2
index_islands <- c(1, 3)
index_nonislands <- c(2, 4)
data <- list(
  list(c(.5, .5, 0, 0, 0, .5), c(.5, 0, 0, .5), c(.5, .5, 0), c(0, 0, .5)), # tip 1
  list(c(0, .5, .5, 1, 1, .5), c(1, .5, 1, .5), c(0, .5, .5), c(1, .5, 1)) # tip 2
  )
get_islandMeanFreqP(index_islands, data, sample_n)
get_nonislandMeanFreqP(index_nonislands, data, sample_n)
get_islandMeanFreqM(index_islands, data, sample_n)
get_nonislandMeanFreqM(index_nonislands, data, sample_n)
```

### Mean SD of observed frequencies per structure type at tree tips

- **Partially-methylated sites**: Let $x_{i,t}$ and $y_{j,t}$ be the frequency of methylation state $p$ in island $i \in\{1, \cdots, I\}$ and non-island $j \in\{1, \cdots, J\}$, respectively. We define the standard deviation of the frequency of $x$ and $y$ at tip $t$ as:

$$
\sigma_{t}(x)=\sqrt{\frac{1}{I-1}\sum_{i=1}^{I}(x_{i,t}-\overline{x_{.,t}})^2} \\
\sigma_{t}(y)=\sqrt{\frac{1}{J-1}\sum_{j=1}^{J}(y_{i,t}-\overline{y_{.,t}})^2}
$$

, where $\overline{x_{.,t}}$ is the mean frequency of state $p$ at tip $t$ in islands and $\overline{y_{.,t}}$ is the mean frequency of state $p$ at tip $t$ in non-islands:

$$
\overline{x_{.,t}}=\frac{1}{I}\sum_{i=1}^{I}x_{i,t} \\
\overline{y_{.,t}}=\frac{1}{J}\sum_{j=1}^{J}x_{j,t}
$$

Therefore, the mean standard deviation across tips is:

$$
\hat{\sigma}(x)=\frac{1}{T}\sum_{t=1}^{T}\sigma_t(x) \\
\hat{\sigma}(y)=\frac{1}{T}\sum_{t=1}^{T}\sigma_t(y) \\
$$

- **Methylated sites**: Let $w_{i,t}$ and $z_{j,t}$ be the frequency of methylation state $m$ in island $i \in\{1, \cdots, I\}$ and non-island $j \in\{1, \cdots, J\}$, respectively. We define the standard deviation of the frequency of $w$ and $z$ at tip $t$ as:

$$
\sigma_{t}(w)=\sqrt{\frac{1}{I-1}\sum_{i=1}^{I}(x_{i,t}-\overline{x_{.,t}})^2} \\
\sigma_{t}(z)=\sqrt{\frac{1}{J-1}\sum_{j=1}^{J}(y_{i,t}-\overline{y_{.,t}})^2}
$$

, where $\overline{x_{.,t}}$ is the mean frequency of state $m$ at tip $t$ in islands and $\overline{z_{.,t}}$ is the mean frequency of state $m$ at tip $t$ in non-islands:

$$
\overline{w_{.,t}}=\frac{1}{I}\sum_{i=1}^{I}w_{i,t} \\
\overline{z_{.,t}}=\frac{1}{J}\sum_{j=1}^{J}z_{j,t}
$$

Therefore, the mean standard deviation across tips is:

$$
\hat{\sigma}(w)=\frac{1}{T}\sum_{t=1}^{T}\sigma_t(w) \\
\hat{\sigma}(z)=\frac{1}{T}\sum_{t=1}^{T}\sigma_t(z) \\
$$
These values can be computed with the functions `get_islandSDFreqP`, `get_nonislandSDFreqP`,`get_islandSDFreqM` and `get_nonislandSDFreqM` as in the following examples:

```{r}
# 1 tip / sample / replicate
sample_n <- 1
index_islands <- c(1, 3)
index_nonislands <- c(2, 4)
data <- list(c(.5, .5, 0, 1, 1, .5), c(.5, 0, 1, .5), c(.5, .5, 0), c(0, 0, .5)) # tip 1
get_islandSDFreqP(index_islands, data, sample_n)
get_nonislandSDFreqP(index_nonislands, data, sample_n)
get_islandSDFreqM(index_islands, data, sample_n)
get_nonislandSDFreqM(index_nonislands, data, sample_n)
```

```{r}
# 2 tip / sample / replicate
sample_n <- 2
index_islands <- c(1, 3)
index_nonislands <- c(2, 4)
data <- list(
  list(c(.5, .5, 0, 0, 0, 1), c(.5, 0, 0, .5), c(1, .5, 0), c(0, 0, .5)), # tip 1
  list(c(0, .5, .5, 1, 1, .5), c(1, .5, 1, .5), c(0, .5, .5), c(1, .5, 1)) # tip 2
  )
get_islandSDFreqP(index_islands, data, sample_n)
get_nonislandSDFreqP(index_nonislands, data, sample_n)
get_islandSDFreqM(index_islands, data, sample_n)
get_nonislandSDFreqM(index_nonislands, data, sample_n)
```

Note that we average across tips as there can be correlations between the mean standard deviation of tips (e.g. under the case of an event affecting the states at two tips).

# Mean correlation of a central segment of methylation states per genomic structure type

Let $s$ be the length of the 2 adjacent segments of central positions to consider.
Let $l$ represent a number of positions to exclude at the start and end of each island/non-island structure.
Define the index positions $p=l+1$ and $p' = s-l$.
Let $x_{i,t,[p,p'-1]}$ and $x_{i,t,[p+1,p']}$ be the methylation states $x \in \{u, p, m\}$ in island $i \in \{1, \cdots ,I\}$ at tip $t$ for the given segments. 
Let $y_{j,t,[p,p'-1]}$ and $y_{j,t,[p+1,p']}$ be the methylation states $y \in \{u, p, m\}$ in non-island $j \in \{1, \cdots ,J\}$ at tip $t$ for the given segments.

We define the mean correlation as:

\[
\overline{Cor(x)_{s,l}}=\frac1I\frac1T\sum_{i=1}^{I}\sum_{t=1}^{T}Cor(x_{i,t,[p:p'-1]}, x_{i,t,[p+1:p']}) \\
\overline{Cor(y)_{s,l}}=\frac1J\frac1T\sum_{j=1}^{J}\sum_{t=1}^{T}Cor(y_{j,t,[p:p'-1]}, y_{j,t,[p+1:p']})
\]

where correlations are computed only for segment pairs where the standard deviation of each segment is non-zero:

\[
\sigma(x_{i,t,[p:p'-1]}) \neq 0 \text{ and } \sigma(x_{i,t,[p+1:p']})\neq 0 \\
\sigma(y_{j,t,[p:p'-1]}) \neq 0 \text{ and } \sigma(y_{j,t,[p+1:p']})\neq 0
\]

In this formulation:

- Only structures with a minimum length of $n = s + 2 \cdot l$ are considered.

- If $l = 0$, all the positions are considered.

These values can be computed with the functions `compute_meanCor_i` for island structures and `compute_meanCor_ni` for non-island structures as in the following examples:

```{r}
# 1 tip / sample / replicate
sample_n <- 1
index_islands <- c(1, 3)
index_nonislands <- c(2, 4) 
data <- list(c(.5, 0, 0, 0, .5, .5, .5, .5, .5, 1, .5, 0, 0, 0, .5, .5, .5, .5, .5, 1, .5, 0, 0, 0, .5, .5, .5, .5, .5, 1), # 30 sites
               c(.5, 1, 1, 1, .5, .5, 1, 1, 1, .5, .5, 1, 1, 1, .5, .5, 1, 1, 1, .5, .5, 1, 1, 1, .5), # 25 sites
               c(.5, 0, 0, .5, 1, .5, 0, 0, .5, 1, .5, 0, 0, .5, 1, .5, 0, 0, .5, 1, .5, 0, 0, .5, 1, .5, 0, 0, .5, 1, .5, 0, 0, .5, 1, .5, 0, 0, .5, 1), # 40 sites
               c(1, 1, 1, .5, .5, .5, 0, 0, 0, .5, 1, 1, 1, .5, .5, .5, 0, 0, 0, .5, 1, 1, 1, .5, .5, .5, 0, 0, 0, .5, .5, 0, 0, 0, .5)) # 35 sites
compute_meanCor_i(index_islands, minN_CpG = 10, shore_length = 5, data, sample_n = 1)
compute_meanCor_ni(index_nonislands, minN_CpG = 10, shore_length = 5, data, sample_n = 1)
```

```{r}
# 2 tip / sample / replicate
sample_n <- 2
index_islands <- c(1, 3)
index_nonislands <- c(2, 4)
data <- list(
  # tip 1
    list(c(.5, 0, 0, 0, .5, .5, .5, .5, .5, 1, .5, 0, 0, 0, .5, .5, .5, .5, .5, 1, .5, 0, 0, 0, .5, .5, .5, .5, .5, 1), # 30 sites
         c(.5, 1, 1, 1, .5, .5, 1, 1, 1, .5, .5, 1, 1, 1, .5, .5, 1, 1, 1, .5, .5, 1, 1, 1, .5), # 25 sites
         c(.5, 0, 0, .5, 1, .5, 0, 0, .5, 1, .5, 0, 0, .5, 1, .5, 0, 0, .5, 1, .5, 0, 0, .5, 1, .5, 0, 0, .5, 1, .5, 0, 0, .5, 1, .5, 0, 0, .5, 1), # 40 sites
         c(1, 1, 1, .5, .5, .5, 0, 0, 0, .5, 1, 1, 1, .5, .5, .5, 0, 0, 0, .5, 1, 1, 1, .5, .5, .5, 0, 0, 0, .5, .5, 0, 0, 0, .5)), # 35 sites
  # tip 2
    list(c(.5, 0, 0, .5, .5, .5, 0, 0, .5, 1, .5, 0, 0, 0, 0, .5, .5, 1, 1, 1, .5, 0, 0, 0, .5, .5, 1, 1, 1, 1), # 30 sites
         c(.5, .5, 1, 1, .5, .5, 1, 1, 1, .5, .5, 0, 0, 0, .5, .5, 1, 1, 1, .5, .5, 1, 1, 1, .5), # 25 sites
         c(.5, 0, 0, .5, 1, .5, 0, 0, .5, 1, .5, 0, 0, .5, .5, .5, 0, 0, .5, 1, 1, 1, 1, .5, 1, .5, 0, 0, .5, 1, .5, 0, 0, .5, 1, .5, 0, 0, .5, 1), # 40 sites
         c(1, 1, 1, .5, .5, .5, 0, 0, 0, .5, 1, 1, 1, 1, .5, .5, 0, 0, 0, .5, 1, 1, 1, .5, .5, .5, .5, .5, 0, .5, .5, .5, .5, 0, .5)) # 35 sites
  )
compute_meanCor_i(index_islands, minN_CpG = 10, shore_length = 5, data, sample_n = 2)
compute_meanCor_ni(index_nonislands, minN_CpG = 10, shore_length = 5, data, sample_n = 2)
```

*Correlation measures how much the state of each site says about the next. So, when there is no variation (e.g. in middle segment state u of a site is followed by state u of the next), there is no correlation.*

## Comparison of tree tips

### Mean Frequency of Methylation State Changes at Single Sites per Structure Type

Let \(S_k\) be the set of sites in the genomic structure indexed by \(k\). For each site \(s \in S_k\), let the methylation state at a given tip \(t\) be denoted as \(m_{k,s,t}\), where \(m_{k,s,t} \in \{0, 0.5, 1\}\) represents the unmethylated, partially-methylated, or methylated state, respectively.

For a given cherry \((t_1, t_2)\), we compute the proportion of sites with differing methylation states between the two tips.

Define the indicator variable:
$$
\delta_{k,s,t_1,t_2} =
\begin{cases}
1, & \text{if } m_{k,s,t_1} \neq m_{k,s,t_2}, \\
0, & \text{if } m_{k,s,t_1} = m_{k,s,t_2}.
\end{cases}
$$

Then, the proportion of sites with different methylation states in structure \(k\) for the cherry \((t_1, t_2)\) is given by:

$$
F_k(t_1, t_2) = \frac{1}{|S_k|} \sum_{s \in S_k} \delta_{k,s,t_1,t_2}
$$

where \(|S_k|\) is the total number of sites in structure \(k\), and the sum counts the number of sites where the methylation state differs between tips \(t_1\) and \(t_2\).

This statistic quantifies the proportion of differing methylation states for a given genomic structure in a cherry.

To compute the weighted mean frequency of methylation changes across all structures in a cherry, we separate island and non-island structures.

For a given cherry \((t_1, t_2)\), let \( \mathcal{I} \) be the set of indices for island structures and \( \mathcal{N} \) be the set of indices for non-island structures. Each structure \( k \) has a total number of sites \( |S_k| \), which we use as weights.

The weighted mean frequency of methylation changes for island structures is:

$$
\bar{F}_{\text{island}}(t_1, t_2) = \frac{\sum_{k \in \mathcal{I}} |S_k| F_k(t_1, t_2)}{\sum_{k \in \mathcal{I}} |S_k|}
$$

Similarly, the weighted mean frequency for non-island structures is:

$$
\bar{F}_{\text{non-island}}(t_1, t_2) = \frac{\sum_{k \in \mathcal{N}} |S_k| F_k(t_1, t_2)}{\sum_{k \in \mathcal{N}} |S_k|}
$$

These expressions compute the mean frequency of methylation state differences across all island and non-island structures, weighted by the number of sites in each structure.

Functions:

- `get_cherryDist(tree)` to get the distances between the tips of each cherry. It identifies each cherry by the tip names and the tip indices. The tip indices correspond to (a) the index from left to right on the newick string, (b) the order of the tip label in the `phylo_object$tip.label`, and (c) the index in the methylation data list `data[[tip]][[structure]]` as obtained with the function `simulate_evolData()` when the given tree has several tips.

- `countSites_cherryMethDiff(cherryDist, data)` to get for each cherry the counts of sites with half-methylation and full-methylation changes per genomic structure (island or non-island).

- `freqSites_cherryMethDiff(tree,data)`. The function first validates the tree structure and extracts pairwise distances between cherry tips with `get_cherryDist`. It then validates the input data and counts half and full methylation differences for each cherry at each structure using `countSites_cherryMethDiff`. Finally, it normalizes these (half and full) counts by the number of sites per structure to compute frequencies. .

- `get_siteFChange_cherry(tree,data)`. Uses `freqSites_cherryMethDiff(tree,data)` and then computes the frequency of sites with any change of methylation state (full or half) for each cherry at each genomic structure.

- `MeanSiteFChange_cherry(tree, data, index_islands, index_nonislands)`. The function computes the per-cherry frequency of sites with different methylation states at each structure (islands and non-islands) using \code{get_siteFChange_cherry}. 
Then it calculates the weighted mean site frequency of methylation changes for each cherry separately for islands and non-islands.

```{r}
# Set example tree and methylation data
tree <- "((a:1.5,b:1.5):2,(c:2,d:2):1.5);"
data <- list(
    list(rep(1,10), rep(0,5), rep(1,8)), # tip a
    list(rep(1,10), rep(0.5,5), rep(0,8)), # tip b
    list(rep(1,10), rep(0.5,5), rep(0,8)), # tip c
    list(c(rep(0,5), rep(0.5, 5)), c(0, 0, 1, 1, 1), c(0.5, 1, rep(0, 6)))) # tip d 
# Set the index for islands and non-island structures
index_islands <- c(2)
index_nonislands <- c(1, 3)

MeanSiteFChange_cherry(data = data, 
                       tree = tree, 
                       index_islands = index_islands, 
                       index_nonislands = index_nonislands)
```

# Fitch

- `get_meanMeth_islands` computes the mean methylation state of all the islands. **Maybe this should allow excluding island borders**