### 2025-03-05 application_examples/abc_inferences/ABC_cichlids.Rmd
Adding ABC_cichlids.Rmd

---
title: "ABC_cichlids"
author: "Sara"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true  # Add a table of contents
    toc_float: false
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Before debug

```{r load params and sumstats}
load("design.RData")
names(sampled_params)[12] <- "TMRCA_smallestCherry"
load("merged_sumStats.RData")
```

```{r set functions to compute expected value and sd of P and M/M+U}
# Set functions to compute expected value and sd of P and M/M+U
    get_expectedValueP <- function(alpha, beta){
      alpha/(alpha+beta)
    }

    get_expectedValueMoverMU <- function(alpha, beta){
      alpha/(alpha+beta)
    }

    get_sdP <- function(alpha, beta){
      var <- (alpha*beta)/((alpha+beta)^2*(alpha+beta+1))
      sqrt(var)
    }

    get_sdMoverMU <- function(alpha, beta){
      var <- (alpha*beta)/((alpha+beta)^2*(alpha+beta+1))
      sqrt(var)
    }
```

```{r filter params corresponding to output files}
# Filter sampled params corresponding to output files
params <- sampled_params[as.integer(rownames(merged_sumStats)),]
print(paste("Number of parameter combinations:", dim(params)[1]))
```

```{r}
# Substitute params alpha and beta for expected value and sd
params_expSD <- params[,c("mu", "iota", "alpha_Ri", "TMRCA_smallestCherry")]
params_expSD$expP_I <- get_expectedValueP(params[,"alpha_pI"], 
                                          params[,"beta_pI"])
params_expSD$expP_NI <- get_expectedValueP(params[,"alpha_pNI"],
                                           params[,"beta_pNI"])   
params_expSD$expMoverMU_I <- get_expectedValueMoverMU(params[,"alpha_mI"],
                                                      params[,"beta_mI"])
params_expSD$expMoverMU_NI <- get_expectedValueMoverMU(params[,"alpha_mNI"],
                                                       params[,"beta_mNI"])
params_expSD$sdP_I <- get_sdP(params[,"alpha_pI"],
                              params[,"beta_pI"])
params_expSD$sdP_NI <- get_sdP(params[,"alpha_pNI"],
                               params[,"beta_pNI"])
params_expSD$sdMoverMU_I <- get_sdMoverMU(params[,"alpha_mI"],
                                          params[,"beta_mI"])
params_expSD$sdMoverMU_NI <- get_sdMoverMU(params[,"alpha_mNI"],
                                           params[,"beta_mNI"])
```

## Params as alpha and beta

```{r, message=FALSE, warning=FALSE}
library(abc)
set.seed(1)
loclinear_cv <- cv4abc(params, merged_sumStats,
                       nval = 100,
                       method = "loclinear",
                       transf = "none",
                       tols = 0.01)
set.seed(1)
ridge_cv <- cv4abc(params, merged_sumStats,
                   nval = 100, 
                   method = "ridge",
                   transf = "none",
                   tols = 0.01)
```

```{r}
loclinear_error <- summary(loclinear_cv)[1,]
ridge_error <- summary(ridge_cv)[1,]
```

```{r}
# Combine into a matrix for grouped bars
error_matrix <- rbind(loclinear_error, ridge_error)

par(mar = c(5, 10, 4, 2))  # Increase left margin for labels
# Create the horizontal bar plot
barplot(error_matrix, 
        horiz = TRUE, 
        beside = TRUE,     # Plot bars side by side
        col = c("red", "steelblue"),  # Red for loclinear, Blue for ridge (cv)
        border = "black",  # Bar border color
        main = "Comparison of Errors",
        xlab = "Error Value",
        names.arg = names(loclinear_error),
        las = 1, 
        cex.names = 0.8,   # Adjust text size
        legend.text = c("Loclinear", "Ridge"),  # Legend
        args.legend = list(x = "topright", bty = "n"))  # Position legend
```

## Params as Expected Value and SD

```{r, message=FALSE, warning=FALSE}
set.seed(1)
loclinear_cv <- cv4abc(params_expSD, merged_sumStats,
                       nval = 100,
                       method = "loclinear",
                       transf = "none",
                       tols = 0.01)
set.seed(1)
ridge_cv <- cv4abc(params_expSD, merged_sumStats,
                   nval = 100, 
                   method = "ridge",
                   transf = "none",
                   tols = 0.01)
```

```{r}
loclinear_error <- summary(loclinear_cv)[1,]
ridge_error <- summary(ridge_cv)[1,]
```

```{r}
# Combine into a matrix for grouped bars
error_matrix <- rbind(loclinear_error, ridge_error)

par(mar = c(5, 10, 4, 2))  # Increase left margin for labels
# Create the horizontal bar plot
barplot(error_matrix, 
        horiz = TRUE, 
        beside = TRUE,     # Plot bars side by side
        col = c("red", "steelblue"),  # Red for loclinear, Blue for ridge (cv)
        border = "black",  # Bar border color
        main = "Comparison of Errors",
        xlab = "Error Value",
        names.arg = names(loclinear_error),
        las = 1, 
        cex.names = 0.8,   # Adjust text size
        legend.text = c("Loclinear", "Ridge"),  # Legend
        args.legend = list(x = "topright", bty = "n"))  # Position legend
```

### Parameter exploration

```{r}
hist(params$mu,
     main = "Frequency of mu in params")
```

```{r}
total_branchLength_unscaled <- 1+1+9+10
total_branchLength_scaled <- params$TMRCA_smallestCherry * total_branchLength_unscaled
expIWE_perIsland <- total_branchLength_scaled * params$mu
hist(expIWE_perIsland,
     main = "Expected Number IWE per Island")
```

```{r}
plot(expIWE_perIsland, merged_sumStats$Fitch_islandGlbSt)
```