### 2025-03-13 extensive_tests/draft-sumstat-vignette.Rmd
Updating vignette with sumstat for mean number of significat freq changes per island and cherry

---
title: "Summary Statistics"
output: 
  rmarkdown::html_vignette:
    toc: true
bibliography: ../vignettes/references.bib
vignette: >
  %\VignetteIndexEntry{Summary Statistics with MethEvolSIM}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, echo=FALSE}
## TODO: Update 
library(devtools)
load_all()
```
# Introduction

Here we present summary statistics that can be used,
e.g. in an Approximate Bayesian Computation method [@csillery2012abc],
to estimate parameters of the methylation dynamics model of MethEvolSIM.
The statistics include mean 
frequencies, standard deviations, neighbor correlations of methylation states in 
island and non-island regions.
Additionally, for analyses comparing different 
samples (comparison of tree tips), some of the summary statistics are based
on the mean per-site frequency of methylation changes and
the minimum number of changes per islands leading to different global methylation
states, calculated with the Fitch algorithm [@fitch1971toward].

The provided set of functions allows the computation of summary statistics for 
methylation data in a genomic region with structural categorization
into two types of structures.
In this  vignette, we will refer to these structures as **islands** and 
**non-island structures**.

The `data` argument must be in one of the following formats, depending on the 
analysis:

1. **For analyses without tip comparison:** A list where each element represents 
the methylation states in a given structure (`data[[structure]]`):

```{r}
# Example: a single sample with 3 genomic structures
# (1) island with 10 partially-methylated sites
# (2) non-island with 5 methylated sites
# (3) island with 15 unmethylated sites

data <- list(rep(0.5, 10),  # Partially methylated
             rep(1,5),      # Methylated
             rep(0,15))     # Unmethylated
data
```

2. **For analyses comparing multiple tips in a phylogenetic tree:** A nested 
list where the first level represents the tips, and the second level represents 
the structures within each tip (`data[[tip]][[structure]]`).

```{r}
# Example: data from 3 tips of a tree,
# each with 3 genomic structures
data <- list(
    # Tip 1
    list(c(rep(0.5,5), rep(0,5)),  # 5 partially methylated, 5 unmethylated
         c(rep(1,4), 0.5),         # 4 methylated, 1 unmethylated
         c(rep(0,7), rep(0.5,8))), # 7 unmethylated, 8 partially methylated
    # Tip 2
    list(c(rep(0.5,9), 1), # 9 partially methylated, 1 methylated
         c(rep(0.5,4), 1), # 4 partially methylated, 1 methylated
         c(rep(0,8), rep(0.5,7))), # 8 unmethylated, 7 partially-methylated
    # Tip 3
    list(c(1, rep(0,8), 1), # first and last methylated, rest unmethylated
         c(rep(0.5,3), rep(1,2)), # 3 methylated, 1 unmethylated
         c(rep(0.5,15)))) # all partially methylated
```

In this case, `data` must be pre-filtered to include only sites present in all 
tips, ensuring a valid comparison between species or samples.

Regardless of the format, methylation values should be represented as follows:

- `0` for unmethylated sites,

- `0.5` for partially methylated sites,

- `1` for methylated sites.

Intermediate values (e.g., obtained from pooled empirical data) should be 
categorized before analysis with `categorize_siteMethSt`.

```{r}
non_categorized_data <- list(
  # Tip 1
    list(c(0.1, 0.7, 0.9), # First structure
         c(0.3, 0.5, 0.9)), # Second structure
  # Tip 2
    list(c(0.2, 0.8, 0.6), # First structure
         c(0.9, 0.4, 0.7)) # Second structure
  )

# Transform the data with custom thresholds
categorized_data <- categorize_siteMethSt(data, u_threshold = 0.15, m_threshold = 0.85)

categorized_data
```

Finally, when phylogenetic data is provided, the order of tips in `data` must 
match the tip order in the Newick tree format (left to right) or the order of 
`tree$tip.label()` when the tree is provided as a `phylo` object from the `ape` 
package.

```{r}
# Example tree in Newick format for the above data
newick_tree <- "((tip1:1, tip2:1):1, tip3:2);"

# Example tree as a phylo object from the ape package
library(ape)
phylo_tree <- read.tree(text = newick_tree)
phylo_tree$tip.label
```

# Summary Statistics

## Frequency of Methylation States per Genomic Structure Type

### Global Mean

- **Partially-methylated sites**: Let $x_{i,t}$ and $y_{j,t}$ be the frequency 
of methylation state $p$ in island $i \in\{1, \cdots, I\}$ and non-island 
$j \in\{1, \cdots, J\}$ at tip $t$, respectively. We define the mean global 
frequency of $x$ and $y$ as:

$$
\overline{x}=\frac1I\sum_{i=1}^{I}\frac1T\sum_{t=1}^{T}x_{i,t} \\
\overline{y}=\frac1J\sum_{j=1}^{J}\frac1T\sum_{t=1}^{T}y_{j,t}
$$

where $x_{i,t}$ and $y_{j,t}$ represent the respective frequencies at each tip.

- **Methylated sites**: Similarly, let $w_{i,t}$ and $z_{j,t}$ be the frequency 
of methylation state $m$ in islands and non-islands. The mean global frequencies 
are:

$$
\overline{w}=\frac1I\sum_{i=1}^{I}\frac1T\sum_{t=1}^{T}w_{i,t} \\
\overline{z}=\frac1J\sum_{j=1}^{J}\frac1T\sum_{t=1}^{T}z_{j,t}
$$

These values can be computed using the following functions:

```{r}
# 1 tip / sample / replicate
sample_n <- 1
index_islands <- c(1, 3)
index_nonislands <- c(2, 4)
data <- list(c(.5, .5, 0, 0, 0, .5), c(.5, 0, 0, .5), c(.5, .5, 0), c(0, 1, .5)) # tip 1
get_islandMeanFreqP(index_islands, data, categorized_data = T, sample_n)
get_nonislandMeanFreqP(index_nonislands, data, categorized_data = T, sample_n)
get_islandMeanFreqM(index_islands, data, categorized_data = T, sample_n)
get_nonislandMeanFreqM(index_nonislands, data, categorized_data = T, sample_n)
```

```{r}
# 2 tip / sample / replicate
sample_n <- 2
index_islands <- c(1, 3)
index_nonislands <- c(2, 4)
data <- list(
  list(c(.5, .5, 0, 0, 0, .5), c(.5, 0, 0, .5), c(.5, .5, 0), c(0, 0, .5)), # tip 1
  list(c(0, .5, .5, 1, 1, .5), c(1, .5, 1, .5), c(0, .5, .5), c(1, .5, 1)) # tip 2
  )
get_islandMeanFreqP(index_islands, data, categorized_data = T, sample_n)
get_nonislandMeanFreqP(index_nonislands, data, categorized_data = T, sample_n)
get_islandMeanFreqM(index_islands, data, categorized_data = T, sample_n)
get_nonislandMeanFreqM(index_nonislands, data, categorized_data = T, sample_n)
```
Note that in the calculation of the means the relative frequencies $x_{i, t}$, 
$y_{j, t}$, $w_{i, t}$ and $z_{j, t}$ are **not** weighted by the lengths of 
island $i$ or non-island $j$.

### Mean SD of Observed Frequencies per Structure Type at Tree Tips

- **Partially-methylated sites**: We define the standard deviation of $x$ and 
$y$ at tip $t$ as:

$$
\sigma_{t}(x)=\sqrt{\frac{1}{I-1}\sum_{i=1}^{I}(x_{i,t}-\overline{x_{.,t}})^2} \\
\sigma_{t}(y)=\sqrt{\frac{1}{J-1}\sum_{j=1}^{J}(y_{j,t}-\overline{y_{.,t}})^2}
$$

where $\overline{x_{.,t}}$ and $\overline{y_{.,t}}$ represent mean frequencies 
at tip $t$. The mean standard deviation across tips is:

$$
\hat{\sigma}(x)=\frac{1}{T}\sum_{t=1}^{T}\sigma_t(x) \\
\hat{\sigma}(y)=\frac{1}{T}\sum_{t=1}^{T}\sigma_t(y)
$$

- **Methylated sites**: Similarly, for $w$ and $z$:

$$
\hat{\sigma}(w)=\frac{1}{T}\sum_{t=1}^{T}\sigma_t(w) \\
\hat{\sigma}(z)=\frac{1}{T}\sum_{t=1}^{T}\sigma_t(z)
$$

These values can be computed using:

```{r}
# 1 tip / sample / replicate
sample_n <- 1
index_islands <- c(1, 3)
index_nonislands <- c(2, 4)
data <- list(c(.5, .5, 0, 1, 1, .5), c(.5, 0, 1, .5), c(.5, .5, 0), c(0, 0, .5))
get_islandSDFreqP(index_islands, data, categorized_data = T, sample_n)
get_nonislandSDFreqP(index_nonislands, data, categorized_data = T, sample_n)
get_islandSDFreqM(index_islands, data, categorized_data = T, sample_n)
get_nonislandSDFreqM(index_nonislands, data, categorized_data = T, sample_n)
```

```{r}
# 2 tip / sample / replicate
sample_n <- 2
index_islands <- c(1, 3)
index_nonislands <- c(2, 4)
data <- list(
  list(c(.5, .5, 0, 0, 0, 1), c(.5, 0, 0, .5), c(1, .5, 0), c(0, 0, .5)), # tip 1
  list(c(0, .5, .5, 1, 1, .5), c(1, .5, 1, .5), c(0, .5, .5), c(1, .5, 1)) # tip 2
  )
get_islandSDFreqP(index_islands, data, categorized_data = T, sample_n)
get_nonislandSDFreqP(index_nonislands, data, categorized_data = T, sample_n)
get_islandSDFreqM(index_islands, data, categorized_data = T, sample_n)
get_nonislandSDFreqM(index_nonislands, data, categorized_data = T, sample_n)
```

Note that we average across tips as there can be correlations between the mean 
standard deviation of tips (e.g. under the case of an event affecting the states 
at two tips).

## Mean Correlation of a Central Segment of Methylation States per Genomic Structure Type

Let $s$ be the length of two adjacent segments of central positions to consider. 
Let $l$ be the number of positions excluded at the start and end of each 
structure. Define:

$$
\overline{Cor(x)_{s,l}}=\frac1I\frac1T\sum_{i=1}^{I}\sum_{t=1}^{T}Cor(x_{i,t,[p:p'-1]}, x_{i,t,[p+1:p']}) \\
\overline{Cor(y)_{s,l}}=\frac1J\frac1T\sum_{j=1}^{J}\sum_{t=1}^{T}Cor(y_{j,t,[p:p'-1]}, y_{j,t,[p+1:p']})
$$

Correlations are computed only for segment pairs where standard deviation is 
non-zero:

$$
\sigma(x_{i,t,[p:p'-1]}) \neq 0 \text{ and } \sigma(x_{i,t,[p+1:p']})\neq 0 \\
\sigma(y_{j,t,[p:p'-1]}) \neq 0 \text{ and } \sigma(y_{j,t,[p+1:p']})\neq 0
$$

In this formulation:

- Only structures with a minimum length of $n = s + 2 \cdot l$ are considered.

- If $l = 0$, all the positions are considered.

These values can be computed with the functions `compute_meanCor_i` for island structures and `compute_meanCor_ni` for non-island structures as in the following examples:

```{r}
# 1 tip / sample / replicate
sample_n <- 1
index_islands <- c(1, 3)
index_nonislands <- c(2, 4) 
data <- list(c(.5, 0, 0, 0, .5, .5, .5, .5, .5, 1, .5, 0, 0, 0, .5, .5, .5, .5, 
               .5, 1, .5, 0, 0, 0, .5, .5, .5, .5, .5, 1), # 30 sites
               c(.5, 1, 1, 1, .5, .5, 1, 1, 1, .5, .5, 1, 1, 1, .5, .5, 1, 1, 1, 
                 .5, .5, 1, 1, 1, .5), # 25 sites
               c(.5, 0, 0, .5, 1, .5, 0, 0, .5, 1, .5, 0, 0, .5, 1, .5, 0, 0, 
                 .5, 1, .5, 0, 0, .5, 1, .5, 0, 0, .5, 1, .5, 0, 0, .5, 1, .5, 
                 0, 0, .5, 1), # 40 sites
               c(1, 1, 1, .5, .5, .5, 0, 0, 0, .5, 1, 1, 1, .5, .5, .5, 0, 0, 0, 
                 .5, 1, 1, 1, .5, .5, .5, 0, 0, 0, .5, 
                 .5, 0, 0, 0, .5)) # 35 sites
compute_meanCor_i(index_islands, minN_CpG = 10, 
                  shore_length = 5, data, sample_n = 1, categorized_data = T)
compute_meanCor_ni(index_nonislands, minN_CpG = 10, 
                   shore_length = 5, data, sample_n = 1, categorized_data = T)
```

```{r}
# 2 tip / sample / replicate
sample_n <- 2
index_islands <- c(1, 3)
index_nonislands <- c(2, 4)
data <- list(
  # tip 1
    list(c(.5, 0, 0, 0, .5, .5, .5, .5, .5, 1, .5, 0, 0, 0, .5, .5, .5, .5, .5, 
           1, .5, 0, 0, 0, .5, .5, .5, .5, .5, 1), # 30 sites
         c(.5, 1, 1, 1, .5, .5, 1, 1, 1, .5, .5, 1, 1, 1, .5, .5, 1, 1, 1, .5, 
           .5, 1, 1, 1, .5), # 25 sites
         c(.5, 0, 0, .5, 1, .5, 0, 0, .5, 1, .5, 0, 0, .5, 1, .5, 0, 0, .5, 1, 
           .5, 0, 0, .5, 1, .5, 0, 0, .5, 1, .5, 0, 0, .5, 1, .5, 0, 0, 
           .5, 1), # 40 sites
         c(1, 1, 1, .5, .5, .5, 0, 0, 0, .5, 1, 1, 1, .5, .5, .5, 0, 0, 0, .5, 
           1, 1, 1, .5, .5, .5, 0, 0, 0, .5, .5, 0, 0, 0, .5)), # 35 sites
  # tip 2
    list(c(.5, 0, 0, .5, .5, .5, 0, 0, .5, 1, .5, 0, 0, 0, 0, .5, .5, 1, 1, 1, 
           .5, 0, 0, 0, .5, .5, 1, 1, 1, 1), # 30 sites
         c(.5, .5, 1, 1, .5, .5, 1, 1, 1, .5, .5, 0, 0, 0, .5, .5, 1, 1, 1, .5, 
           .5, 1, 1, 1, .5), # 25 sites
         c(.5, 0, 0, .5, 1, .5, 0, 0, .5, 1, .5, 0, 0, .5, .5, .5, 0, 0, .5, 1, 
           1, 1, 1, .5, 1, .5, 0, 0, .5, 1, .5, 0, 0, .5, 1, .5, 0, 0, 
           .5, 1), # 40 sites
         c(1, 1, 1, .5, .5, .5, 0, 0, 0, .5, 1, 1, 1, 1, .5, .5, 0, 0, 0, .5, 1, 
           1, 1, .5, .5, .5, .5, .5, 0, .5, .5, .5, .5, 0, .5)) # 35 sites
  )
compute_meanCor_i(index_islands, minN_CpG = 10, 
                  shore_length = 5, data, sample_n = 2, categorized_data = T)
compute_meanCor_ni(index_nonislands, minN_CpG = 10, 
                   shore_length = 5, data, sample_n = 2, categorized_data = T)
```

Note that correlation measures how much the state of each site says about the 
next. So, when there is no variation (e.g. in middle segment state u of a site 
is followed by state u of the next), there is no correlation.

## Comparison of tree tips

### Mean Frequency of Methylation State Changes at Single Sites per Structure Type

Let $S_k$ be the set of sites in the genomic structure indexed by $k$. 
For each site $s \in S_k$, let the methylation state at a given tip $t$ be 
denoted as $m_{k,s,t}$, where $m_{k,s,t} \in \{0, 0.5, 1\}$ represents the 
unmethylated, partially-methylated, or methylated state, respectively.

For a given cherry $(t_1, t_2)$, that is two tips $t_1$ and $t_2$ that are direct
offspring of the same internal node, we compute the proportion of sites with 
differing methylation states between the two tips.

Define the indicator variable:
$$
\delta_{k,s,t_1,t_2} =
\begin{cases}
1, & \text{if } m_{k,s,t_1} \neq m_{k,s,t_2}, \\
0, & \text{if } m_{k,s,t_1} = m_{k,s,t_2}.
\end{cases}
$$

Then, the proportion of sites with different methylation states in structure $k$ 
for the cherry $(t_1, t_2)$ is given by:

$$
F_k(t_1, t_2) = \frac{1}{|S_k|} \sum_{s \in S_k} \delta_{k,s,t_1,t_2}
$$

where $|S_k|$ is the total number of sites in structure \(k\), and the sum 
counts the number of sites where the methylation state differs between tips 
$t_1$ and $t_2$.

This statistic quantifies the proportion of differing methylation states for a 
given genomic structure in a cherry.

To compute the weighted mean frequency of methylation changes across all 
structures in a cherry, we separate island and non-island structures.

For a given cherry $(t_1, t_2)$, let $\mathcal{I}$ be the set of indices 
for island structures and $\mathcal{N}$ be the set of indices for non-island 
structures. Each structure $k$ has a total number of sites $|S_k|$,
which we use as weights.

The weighted mean frequency of methylation changes for island structures is:

$$
\bar{F}_{\text{island}}(t_1, t_2) = \frac{\sum_{k \in \mathcal{I}} |S_k| F_k(t_1, t_2)}{\sum_{k \in \mathcal{I}} |S_k|}
$$

Similarly, the weighted mean frequency for non-island structures is:

$$
\bar{F}_{\text{non-island}}(t_1, t_2) = \frac{\sum_{k \in \mathcal{N}} |S_k| F_k(t_1, t_2)}{\sum_{k \in \mathcal{N}} |S_k|}
$$

These expressions compute the mean frequency of methylation state differences 
across all island and non-island structures, weighted by the number of sites in 
each structure.

Functions:

- `get_cherryDist(tree)` to get the distances between the tips of each cherry. 
It identifies each cherry by the tip names and the tip indices. 
The tip indices correspond to (a) the index from left to right on the newick 
string, (b) the order of the tip label in the `phylo_object$tip.label`, and (c) 
the index in the methylation data list `data[[tip]][[structure]]` as obtained 
with the function `simulate_evolData()` when the given tree has several tips.

- `countSites_cherryMethDiff(cherryDist, data)` to get for each cherry the 
counts of sites with half-methylation and full-methylation changes per genomic 
structure (island or non-island).

- `freqSites_cherryMethDiff(tree,data)`. The function first validates the tree 
structure and extracts pairwise distances between cherry tips 
with `get_cherryDist`. It then validates the input data and counts half and full 
methylation differences for each cherry at each structure using 
`countSites_cherryMethDiff`. Finally, it normalizes these (half and full) counts 
by the number of sites per structure to compute frequencies.

- `get_siteFChange_cherry(tree,data)` uses `freqSites_cherryMethDiff(tree,data)` 
and then computes the frequency of sites with any change of methylation state 
(full or half) for each cherry at each genomic structure.

- `MeanSiteFChange_cherry(tree, data, index_islands, index_nonislands)`. The 
function computes the per-cherry frequency of sites with different methylation 
states at each structure (islands and non-islands) using `get_siteFChange_cherry`. 
Then it calculates the weighted mean site frequency of methylation changes for 
each cherry separately for islands and non-islands.

```{r}
# Set example tree and methylation data
tree <- "((a:1.5,b:1.5):2,(c:2,d:2):1.5);"
data <- list(
    list(rep(1,10), rep(0,5), rep(1,8)), # tip a
    list(rep(1,10), rep(0.5,5), rep(0,8)), # tip b
    list(rep(1,10), rep(0.5,5), rep(0,8)), # tip c
    list(c(rep(0,5), rep(0.5, 5)), c(0, 0, 1, 1, 1), c(0.5, 1, rep(0, 6)))) # d

# Set the index for islands and non-island structures
index_islands <- c(2)
index_nonislands <- c(1, 3)

MeanSiteFChange_cherry(data = data, 
                       categorized_data = T, 
                       tree = tree, 
                       index_islands = index_islands, 
                       index_nonislands = index_nonislands)
```

### Fitch Parsimony Estimate for Minimum Global Methylation State Changes at CpG Islands

Let $I$ be the set of CpG islands in a given genomic region For each island 
$i \in I$, the mean methylation level across sites is computed with 
`get_meanMeth_islands(index_islands, data)` and categorized into one of three 
global methylation states—unmethylated ($u$), partially methylated ($p$), or 
methylated ($m$)—based on user-defined thresholds $u_{\text{thresh}}$ and 
$m_{\text{thresh}}$. This categorization is given by the function:

$$
\operatorname{categorize\_islandGlbSt}(\mu_i, u_{\text{thresh}}, m_{\text{thresh}}) =
\begin{cases}
  u, & \text{if } \mu_i \leq u_{\text{thresh}}, \\
  m, & \text{if } \mu_i \geq m_{\text{thresh}}, \\
  p, & \text{otherwise}.
\end{cases}
$$

Given a phylogenetic tree $T$ with tips corresponding to sampled species or 
individuals, we aim to estimate the minimum number of state changes required to 
explain the observed distribution of island methylation states across the tips. 
This is achieved using the Fitch algorithm [@fitch1971toward] as
implemented in `compute_fitch(meth, tree)`.

The function `computeFitch_islandGlbSt(T, M)`, where $M$ is the matrix of 
categorized methylation states across tips, returns a vector $C$ where each entry 
$C_i$ represents the minimum number of changes required for island $i$ under the 
Fitch parsimony criterion.

```{r}
# Example with data from a single island structure 
# and three tips

tree <- "((bla:1,bah:1):2,booh:2);"

data <- list(
  #Tip 1
  list(c(rep(1,9), rep(0,1))), # m
  #Tip 2
  list(c(rep(0.5,10))), # p
  #Tip 3
  list(c(rep(0.5,9), rep(0.5,1)))) # p

index_islands <- c(1)

computeFitch_islandGlbSt(index_islands, data, tree, 
                         u_threshold = 0.1, m_threshold = 0.9)
```
```{r}
# Example: data from a genomic region consisting on 3 structures with 10 sites each
# one island, one non-island, one island
# and a tree with 8 tips
tree <- "(((a:1,b:1):1,(c:1,d:1):1):1,((e:1,f:1):1,(g:1,h:1):1):1);"
data <- list(
  #Tip 1
  list(c(rep(1,5), rep(0,5)), # p
       c(rep(0,9), 1), 
        c(rep(1,8), rep(0.5,2))), # m
  #Tip 2
  list(c(rep(0.5,9), rep(0.5,1)), # p
        c(rep(0.5,9), 1), 
       c(rep(0,8), rep(0.5,2))), # u
  #Tip 3
  list(c(rep(1,9), rep(0.5,1)), # m
       c(rep(0.5,9), 1), 
       c(rep(0.5,8), rep(0.5,2))), # p
  #Tip 4
  list(c(rep(1,9), rep(0.5,1)), # m
       c(rep(1,9), 0), 
       c(rep(0.5,8), rep(0.5,2))), # p
  #Tip 5
  list(c(rep(0,5), rep(0,5)), # u
       c(rep(0,9), 1), 
       c(rep(0.5,8), rep(0.5,2))), # p
  #Tip 6
  list(c(rep(0,9), rep(0.5,1)), # u
       c(rep(0.5,9), 1), 
       c(rep(1,8), rep(0.5,2))), # m
  #Tip 7
  list(c(rep(0,9), rep(0.5,1)), # u
       c(rep(0.5,9), 1), 
       c(rep(0,8), rep(0.5,2))), # u
  #Tip 8
  list(c(rep(0,9), rep(0.5,1)), # u
       c(rep(1,9), 0), 
       c(rep(0,9), rep(0.5,1)))) # u

index_islands <- c(1,3)

computeFitch_islandGlbSt(index_islands, data, tree, 
                         u_threshold = 0.1, m_threshold = 0.9)
```

For genomic regions containing multiple CpG islands, we can summarize the 
per-island estimates using the mean to obtain an overall measure of the minimum 
state changes.

```{r}
mean(computeFitch_islandGlbSt(index_islands, data, tree, 
                              u_threshold = 0.1, m_threshold = 0.9))
```

### Mean Number of Significant Methylation Frequency Changes per Island and Cherry

A cherry $c$ is defined as a pair of direct offspring (tips \( t_1 \) and \( t_2 \)) 
of the same internal node in the phylogenetic tree. 
For a given cherry $c$, let $\mathcal{I}$ be the set of indices for island structures. 
For each island structure $i \in \mathcal{I}$ at each cherry tip $t_c \in {t_{c,1}, t_{c,2}$ in each phylogenetic tree cherries, 
the number of sites in each 
methylation state (unmethylated \( 0 \), partially-methylated \( 0.5 \), 
methylated \( 1 \)) is counted. This count is represented as:

$$
\text{count}_{\text{UPM}}(c,t_c, i) = \left( n_0, n_{0.5}, n_1 \right)
$$

where \( n_0\), \( n_{0.5}\), and \( n_1\) are the counts of unmethylated, 
partially-methylated, and methylated sites at tip \( t_c\), respectively.

The distribution of methylation states at each island between the two tips \( t_1 \) 
and \( t_2 \) of a given cherry is compared using a chi-squared test. 
The null hypothesis \( H_0 \) is that the distributions of methylation states at 
the two tips are the same. The test statistic is calculated using the contingency 
table \( T \) of the methylation state counts for each tip:

$$
H_0: \text{distribution}(t_{c,1}) = \text{distribution}(t_{c,2})
$$

The contingency table \( T_{c,i} \) for tips \( t_{c,1} \) and \( t_{c,2} \) is:

$$
T_{c,i} = \begin{pmatrix}
n_0^{t_{c,1}} & n_{0.5}^{t_{c,1}} & n_1^{t_{c,1}} \\
n_0^{t_{c,2}} & n_{0.5}^{t_{c,2}} & n_1^{t_{c,2}}
\end{pmatrix}
$$

The p-value for the chi-squared test is calculated via Monte Carlo simulation to 
improve reliability, even when the expected frequencies do not meet the assumptions 
of the chi-squared test (i.e., expected counts of at least 5 in each category).

For each cherry $c$ and each island $i$, the methylation data at each cherry tip 
for the specified genomic structures (islands) are compared using the chi-squared test.

$$
p_{\text{value}}(c,i) = \chi^2\text{-test}(T_{c,i})
$$

These p-values are stored in a dataframe for each cherry and for each island.

The significance of the methylation frequency changes is determined based on a 
user-defined threshold \( p_{\text{threshold}} \). If the p-value is smaller than 
the threshold, the change is considered significant. This is done for each cherry 
and each island \( i \in \mathcal{I} \) by comparing each p-value to the threshold:

$$
\text{Significant Change}_{c,i} = 
\begin{cases} 
1, & \text{if } p_{\text{value}}(c,i) < p_{\text{threshold}} \\
0, & \text{if } p_{\text{value}}(c,i) \ge p_{\text{threshold}}
\end{cases}
$$

For each cherry , the mean number of significant changes across all islands 
\( i \in \mathcal{I} \) is computed as the average of the significant changes for each island:

$$
\text{Mean Number of Significant Changes per Island}_{c} = \frac{1}{|\mathcal{I}|} \sum_{i \in \mathcal{I}} \text{Significant Change}_{c,i}
$$

where \( |\mathcal{I}| \) is the total number of islands.

Functions:

- `compare_CherryFreqs(tip1, tip2)` to perform a chi-squared test to compare the
distribution of methylation states between two cherry tips.

- `pValue_CherryFreqsChange_i(data, index_islands, tree)` uses `compare_CherryFreqs(tip1, tip2)` to get for each cherry 
and each island the chi-square p-value of the comparison of the distribution of
methylation states between two cherry tips.

- `mean_CherryFreqsChange_i(data, categorized_data = FALSE, index_islands, tree, pValue_threshold)`. The function
uses `pValue_CherryFreqsChange_i(data, index_islands, tree)` and counts the mean
number of significant changes per island at each cherry.

```{r}
# Set example tree and methylation data
  tree <- "((a:1,b:1):2,c:2);"
  data <- list(
    #Tip a
    list(c(rep(1,9), rep(0,1)), # Structure 1: island
         c(rep(0,9), 1), # Structure 2: non-island
         c(rep(0,9), rep(0.5,1))),  # Structure 3: island
    #Tip b
    list(c(rep(0,9), rep(0.5,1)),  # Structure 1: island
         c(rep(0.5,9), 1), # Structure 2: non-island
         c(rep(0,9), rep(0,1))), # Structure 3: island
    #Tip c
    list(c(rep(1,9), rep(0.5,1)),  # Structure 1: island
         c(rep(0.5,9), 1), # Structure 2: non-island
         c(rep(0,9), rep(0.5,1)))) # Structure 3: island

  index_islands <- c(1,3)

  mean_CherryFreqsChange_i(data, categorized_data = T,
                           index_islands, tree,
                           pValue_threshold = 0.05)
```

# References